Manuscript

### 2018-10-29 manuscript.Rmd
work on G2P summ plots

```{r}
library(bsselectR)
library(data.table)
library(ggpubr)
library(ggsignif)
library(kableExtra)
library(reshape2)
library(tidyverse)
```

@@

crispr_data_signif_g2p_gene_plot <- ggplot(data = crispr_data_signif_g2p_gene, aes(x = Mutation_Status_Nonsilent, y = Score)) +
  facet_wrap(~ Hugo_Symbol, drop = FALSE, nrow = 2, scales = "free_y") +
  geom_jitter(aes(color = Mutation_Status_Nonsilent), alpha = 0.5, width = 0.3) +
  geom_boxplot(color = "black", fill = NA, outlier.shape = NA) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  scale_color_manual(values = crispr_data_signif_g2p_gene_color) + 
  geom_text(data = crispr_data_signif_g2p_gene_text, mapping = aes(x = 1.5, y = Text_y, label = p.signif), nudge_y = 0.2) +
  theme(legend.position = "top") +
  labs(y = "CERES Score",
       x = "Lineage",
       color = "Mutation Status",
       title = "CERES scores for select Level A G2P genes",
       subtitle = "Genes are sorted by increasing p-value.")
crispr_data_signif_g2p_gene_plot
# Genes with adjusted p-values < 0.3 are shown.\nBH-adjusted p-values: * p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001\n
# ggsave("./plots_18Q3/manuscript/crispr_data_signif_g2p_gene_plot.png", crispr_data_signif_g2p_gene_plot, width = 12, height = 7, units = "in")
```

@@

crispr_data_signif_g2p_lineage_plot <- ggplot(data = crispr_data_signif_g2p_lineage, aes(x = group_general_lineage_name, y = Score)) +
  facet_wrap(~ Hugo_Symbol, nrow = 4, scales = "free_y") +
  geom_point(aes(color = Mutation_Status_Nonsilent), alpha = 0.5) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  scale_color_manual(values = crispr_data_signif_g2p_lineage_color) +
  geom_text(data = crispr_data_signif_g2p_lineage_text, mapping = aes(x = group_general_lineage_name, y = Text_y, label = p.signif), angle = 90, nudge_y = 0.1) +
  theme(legend.position = "top", axis.text.x = element_text(angle = 90, hjust = 1, vjust = 1, size = 10)) +
  labs(y = "CERES Score",
       x = "Lineage",
       color = "Mutation Status",
       title = "CERES score by cell line lineage for significant Level A G2P genes",
       subtitle = "* p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001\nGenes are sorted by increasing lineage-agnostic p-value.")
crispr_data_signif_g2p_lineage_plot

@@

```{r fig.height=4, fig.width=8}
g2p_druggene_p_summ <- g2p %>% group_by(Drug_Gene) %>% tally()
g2p_druggene_p_summ <- merge(g2p_druggene_p_summ, ccle_signif_g2p[, c("Drug_Gene", "p")], by = "Drug_Gene", all.x = TRUE)
g2p_druggene_p_summ <- merge(g2p_druggene_p_summ, ctrp_signif_g2p[, c("Drug_Gene", "p")], by = "Drug_Gene", all.x = TRUE, suffixes = c("_CCLE", "_CTRP"))
g2p_druggene_p_summ <- merge(g2p_druggene_p_summ, gdsc_signif_g2p[, c("Drug_Gene", "p")], by = "Drug_Gene", all.x = TRUE)
colnames(g2p_druggene_p_summ)[colnames(g2p_druggene_p_summ) == "p_CCLE"] <- "CCLE"
colnames(g2p_druggene_p_summ)[colnames(g2p_druggene_p_summ) == "p_CTRP"] <- "CTRP"
colnames(g2p_druggene_p_summ)[colnames(g2p_druggene_p_summ) == "p"] <- "GDSC"
g2p_druggene_p_summ_melt <- melt(g2p_druggene_p_summ, id.vars = c("Drug_Gene", "n"), measure.vars = c("CCLE", "CTRP", "GDSC"), variable.name = "Dataset", value.name = "p")

g2p_druggene_p_summ_plot <- ggplot(data = g2p_druggene_p_summ_melt, aes(x = p, y = n, shape = Dataset, color = Dataset)) +
  geom_point(size = 2) +
  geom_smooth(method = "lm", se = FALSE, lwd = 0.5) +
  scale_color_manual(values = c("green3", "darkslategray4", "darkturquoise")) +
  labs(x = "Gene-drug association p-value by mutation status", y = "Number of level A G2P associations")
g2p_druggene_p_summ_plot
# ggsave("./plots_18Q3/manuscript/g2p_druggene_p_summ_plot", g2p_druggene_p_summ_plot.png, device = "png", dpi = 450, width = 8, height = 4, units = "in")
```

>>>>

```{r}
library(bsselectR)
library(data.table)
library(ggExtra)
library(ggpubr)
library(ggsignif)
library(kableExtra)
library(reshape2)
library(tidyverse)
```

@@

crispr_data_signif_g2p_gene_plot <- ggplot(data = crispr_data_signif_g2p_gene, aes(x = Mutation_Status_Nonsilent, y = Score)) +
  facet_wrap(~ Hugo_Symbol, drop = FALSE, nrow = 2, scales = "free_y") +
  geom_jitter(aes(color = Mutation_Status_Nonsilent), alpha = 0.5, width = 0.3) +
  geom_boxplot(color = "black", fill = NA, outlier.shape = NA) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  scale_color_manual(values = crispr_data_signif_g2p_gene_color) + 
  geom_text(data = crispr_data_signif_g2p_gene_text, mapping = aes(x = 1.5, y = Text_y, label = p.signif), nudge_y = 0.2) +
  theme(legend.position = "top", axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.title.x = element_blank()) +
  labs(y = "CERES Score",
       color = "Mutation Status",
       title = "CERES scores for select Level A G2P genes",
       subtitle = "Genes are sorted by increasing p-value.")
crispr_data_signif_g2p_gene_plot

# ggsave("./plots_18Q3/manuscript/crispr_data_signif_g2p_gene_plot.png", crispr_data_signif_g2p_gene_plot, width = 12, height = 7, units = "in")
```

@@

crispr_data_signif_g2p_lineage_plot <- ggplot(data = crispr_data_signif_g2p_lineage, aes(x = group_general_lineage_name, y = Score)) +
  facet_wrap(~ Hugo_Symbol, nrow = 4, scales = "free_y") +
  geom_point(aes(color = Mutation_Status_Nonsilent), alpha = 0.5) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  scale_color_manual(values = crispr_data_signif_g2p_lineage_color) +
  geom_text(data = crispr_data_signif_g2p_lineage_text, mapping = aes(x = group_general_lineage_name, y = Text_y, label = p.signif), angle = 90, vjust = 0.8, nudge_y = 0.1) +
  theme(legend.position = "top", axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4, size = 10)) +
  labs(y = "CERES Score",
       x = "Lineage",
       color = "Mutation Status",
       title = "CERES score by cell line lineage for significant Level A G2P genes",
       subtitle = "* p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001\nGenes are sorted by increasing lineage-agnostic p-value.")
crispr_data_signif_g2p_lineage_plot

@@

```{r fig.height=4, fig.width=8}
g2p_druggene_p_tally <- g2p %>% group_by(Drug_Gene) %>% tally()
g2p_druggene_p_summ <- merge(g2p_druggene_p_tally, ccle_signif_g2p[, c("Drug_Gene", "p")], by = "Drug_Gene", all.x = TRUE)
g2p_druggene_p_summ <- merge(g2p_druggene_p_summ, ctrp_signif_g2p[, c("Drug_Gene", "p")], by = "Drug_Gene", all.x = TRUE, suffixes = c("_CCLE", "_CTRP"))
g2p_druggene_p_summ <- merge(g2p_druggene_p_summ, gdsc_signif_g2p[, c("Drug_Gene", "p")], by = "Drug_Gene", all.x = TRUE)
colnames(g2p_druggene_p_summ)[colnames(g2p_druggene_p_summ) == "p_CCLE"] <- "CCLE"
colnames(g2p_druggene_p_summ)[colnames(g2p_druggene_p_summ) == "p_CTRP"] <- "CTRP"
colnames(g2p_druggene_p_summ)[colnames(g2p_druggene_p_summ) == "p"] <- "GDSC"
g2p_druggene_p_summ_melt <- melt(g2p_druggene_p_summ, id.vars = c("Drug_Gene", "n"), measure.vars = c("CCLE", "CTRP", "GDSC"), variable.name = "Dataset", value.name = "p")

g2p_druggene_p_summ_plot <- ggplot(data = g2p_druggene_p_summ_melt, aes(x = p, y = n, shape = Dataset, color = Dataset)) +
  geom_point(size = 2) +
  scale_color_manual(values = c("green3", "darkslategray4", "darkturquoise")) +
  labs(x = "Gene-drug association p-value by mutation status", y = "Number of level A G2P associations")
g2p_druggene_p_summ_plot
# ggsave("./plots_18Q3/manuscript/g2p_druggene_p_summ_plot.png", g2p_druggene_p_summ_plot, device = "png", dpi = 450, width = 8, height = 4, units = "in")
```

```{r}
ccle_druggene_mad_summ <- ccle_data_ptmuts %>% group_by(Drug_Gene, Mutation_Status_Nonsilent) %>% summarize(Mean_AUC = mean(AUC))
ccle_druggene_mad_summ <- dcast(ccle_druggene_mad_summ, Drug_Gene ~ Mutation_Status_Nonsilent)
ccle_druggene_mad_summ$MAD <- abs(ccle_druggene_mad_summ$Mutant - ccle_druggene_mad_summ$Other)
ccle_druggene_p_summ <- ccle_signif_g2p_ByDrug_all
ccle_druggene_p_summ$Drug_Gene <- paste0(ccle_druggene_p_summ$Drug, "_", ccle_druggene_p_summ$Hugo_Symbol)
ccle_druggene_summ <- merge(merge(g2p_druggene_p_tally, ccle_druggene_p_summ[, c("Drug_Gene", "p")], by = "Drug_Gene", all.y = TRUE), ccle_druggene_mad_summ, by = "Drug_Gene", all = TRUE)
ccle_druggene_summ$n <- ifelse(is.na(ccle_druggene_summ$n), 0, ccle_druggene_summ$n)
ccle_druggene_summ$InG2P <- ifelse(ccle_druggene_summ$n != 0, "Yes", "No")
ccle_druggene_summ$Dataset <- "CCLE"
ccle_p_ttest <- t.test(ccle_druggene_summ$p ~ ccle_druggene_summ$InG2P)
ccle_p_ttest_res <- paste0("t = ", round(ccle_p_ttest$statistic, 3), ", df = ", round(ccle_p_ttest$parameter, 3), ", p = ", round(ccle_p_ttest$p.value, 3))
ccle_mad_ttest <- t.test(ccle_druggene_summ$MAD ~ ccle_druggene_summ$InG2P)
ccle_mad_ttest_res <- paste0("t = ", round(ccle_mad_ttest$statistic, 3), ", df = ", round(ccle_mad_ttest$parameter, 3), ", p = ", round(ccle_mad_ttest$p.value, 3))

ctrp_druggene_mad_summ <- ctrp_data_ptmuts %>% group_by(Drug_Gene, Mutation_Status_Nonsilent) %>% summarize(Mean_AUC = mean(AUC))
ctrp_druggene_mad_summ <- dcast(ctrp_druggene_mad_summ, Drug_Gene ~ Mutation_Status_Nonsilent)
ctrp_druggene_mad_summ$MAD <- abs(ctrp_druggene_mad_summ$Mutant - ctrp_druggene_mad_summ$Other)
ctrp_druggene_p_summ <- ctrp_signif_g2p_ByDrug_all
ctrp_druggene_p_summ$Drug_Gene <- paste0(ctrp_druggene_p_summ$Drug, "_", ctrp_druggene_p_summ$Hugo_Symbol)
ctrp_druggene_summ <- merge(merge(g2p_druggene_p_tally, ctrp_druggene_p_summ[, c("Drug_Gene", "p")], by = "Drug_Gene", all.y = TRUE), ctrp_druggene_mad_summ, by = "Drug_Gene", all = TRUE)
ctrp_druggene_summ$n <- ifelse(is.na(ctrp_druggene_summ$n), 0, ctrp_druggene_summ$n)
ctrp_druggene_summ$InG2P <- ifelse(ctrp_druggene_summ$n != 0, "Yes", "No")
ctrp_druggene_summ$Dataset <- "CTRP"
ctrp_p_ttest <- t.test(ctrp_druggene_summ$p ~ ctrp_druggene_summ$InG2P)
ctrp_p_ttest_res <- paste0("t = ", round(ctrp_p_ttest$statistic, 3), ", df = ", round(ctrp_p_ttest$parameter, 3), ", p = ", round(ctrp_p_ttest$p.value, 3))
ctrp_mad_ttest <- t.test(ctrp_druggene_summ$MAD ~ ctrp_druggene_summ$InG2P)
ctrp_mad_ttest_res <- paste0("t = ", round(ctrp_mad_ttest$statistic, 3), ", df = ", round(ctrp_mad_ttest$parameter, 3), ", p = ", round(ctrp_mad_ttest$p.value, 3))

gdsc_druggene_mad_summ <- gdsc_data_ptmuts %>% group_by(Drug_Gene, Mutation_Status_Nonsilent) %>% summarize(Mean_AUC = mean(AUC))
gdsc_druggene_mad_summ <- dcast(gdsc_druggene_mad_summ, Drug_Gene ~ Mutation_Status_Nonsilent)
gdsc_druggene_mad_summ$MAD <- abs(gdsc_druggene_mad_summ$Mutant - gdsc_druggene_mad_summ$Other)
gdsc_druggene_p_summ <- gdsc_signif_g2p_ByDrug_all
gdsc_druggene_p_summ$Drug_Gene <- paste0(gdsc_druggene_p_summ$Drug, "_", gdsc_druggene_p_summ$Hugo_Symbol)
gdsc_druggene_summ <- merge(merge(g2p_druggene_p_tally, gdsc_druggene_p_summ[, c("Drug_Gene", "p")], by = "Drug_Gene", all.y = TRUE), gdsc_druggene_mad_summ, by = "Drug_Gene", all = TRUE)
gdsc_druggene_summ$n <- ifelse(is.na(gdsc_druggene_summ$n), 0, gdsc_druggene_summ$n)
gdsc_druggene_summ$InG2P <- ifelse(gdsc_druggene_summ$n != 0, "Yes", "No")
gdsc_druggene_summ$Dataset <- "GDSC"
gdsc_p_ttest <- t.test(gdsc_druggene_summ$p ~ gdsc_druggene_summ$InG2P)
gdsc_p_ttest_res <- paste0("t = ", round(gdsc_p_ttest$statistic, 3), ", df = ", round(gdsc_p_ttest$parameter, 3), ", p = ", round(gdsc_p_ttest$p.value, 3))
gdsc_mad_ttest <- t.test(gdsc_druggene_summ$MAD ~ gdsc_druggene_summ$InG2P)
gdsc_mad_ttest_res <- paste0("t = ", round(gdsc_mad_ttest$statistic, 3), ", df = ", round(gdsc_mad_ttest$parameter, 3), ", p = ", round(gdsc_mad_ttest$p.value, 3))
```

```{r fig.height=4, fig.width=12}
g2p_druggene_all_summ <- rbind(ccle_druggene_summ, ctrp_druggene_summ, gdsc_druggene_summ)
g2p_druggene_all_summ_text <- data.frame(Dataset = c("CCLE", "CTRP", "GDSC"), 
                                         Test_Results_p = c(ccle_p_ttest_res, ctrp_p_ttest_res, gdsc_p_ttest_res),
                                         Test_Results_MAD = c(ccle_mad_ttest_res, ctrp_mad_ttest_res, gdsc_mad_ttest_res))

g2p_druggene_all_mad_summ_plot <- ggplot(data = g2p_druggene_all_summ) +
  facet_wrap(~ Dataset, scales = "free_x") +
  geom_point(mapping = aes(color = InG2P, x = MAD, y = n), alpha = 0.5) +
  scale_y_continuous(breaks = seq(0, 110, by = 10), labels = seq(0, 110, by = 10)) +
  geom_text(data = g2p_druggene_all_summ_text, mapping = aes(x = c(0.75, 4, 0.2), y = 110, label = Test_Results_MAD)) +
  labs(x = "Gene-drug association difference in mean AUC between mutation groups", y = "Number of level A G2P associations", color = "Does the gene-drug combination have level A G2P indication?") +
  geom_rug(size = 1) +
  theme(legend.position = "top")
g2p_druggene_all_mad_summ_plot

g2p_druggene_all_p_summ_plot <- ggplot(data = g2p_druggene_all_summ) +
  facet_wrap(~ Dataset, scales = "free_x") +
  geom_point(mapping = aes(color = InG2P, x = p, y = n), alpha = 0.5) +
  scale_y_continuous(breaks = seq(0, 110, by = 10), labels = seq(0, 110, by = 10)) +
  scale_x_continuous(breaks = seq(0, 1, by = 0.2), labels = seq(0, 1, by = 0.2)) +
  geom_text(data = g2p_druggene_all_summ_text, mapping = aes(x = 0.5, y = 110, label = Test_Results_p)) +
  labs(x = "Gene-drug association p-value by mutation status", y = "Number of level A G2P associations", color = "Does the gene-drug combination have level A G2P indication?") +
  theme(legend.position = "top")
g2p_druggene_all_p_summ_plot

# ggsave("./plots_18Q3/manuscript/g2p_druggene_all_mad_summ_plot.png", g2p_druggene_all_mad_summ_plot, device = "png", dpi = 450, width = 12, height = 4, units = "in")
# ggsave("./plots_18Q3/manuscript/g2p_druggene_all_p_summ_plot.png", g2p_druggene_all_p_summ_plot, device = "png", dpi = 450, width = 12, height = 4, units = "in")
```
